---
title: "Exploring the Nepal Earthquake 2015 data for prediction and analysis"
author: "Team HCI"
date: "March 3, 2016"
output: html_document
---
  
##Introduction
  
  This project involves exploring Nepal Earthquake 2015 data for prediction and analysis database. This database tracks the total death, injury, damage done by the earthquake by zonal regions. 

## Loading the data

### Load required libraries
```{r echo=TRUE, tidy=TRUE, warning=FALSE, message=FALSE}


affectedFig= read.csv("affected figures.csv", na = "NA")



#get aggregate total affected fig for each zone
totdeathbyzone <- aggregate(Tot_Deaths ~ ZONE, affectedFig, sum)
totdeathbyzone <- totdeathbyzone[(totdeathbyzone$Tot_Deaths  !=0),]
#tothousehold <- aggregate(Total.Household ~ ZONE, affectedFig, sum)

tothousehold <- as.numeric(affectedFig$Total.Household)

#totpopn <- aggregate(Total.Population ~ ZONE, affectedFig, sum)


#get aggregate total affected fig for each zone
#totinjured <- aggregate(Total_Injured ~ ZONE, affectedFig, sum)

#open graphic device for plotting 
#png("plot3.png")#, width=480, height=480)


#dev.off() #close the graphical device


```

### Summary of Data
Take a quick look at the data available. 

```{r echo=TRUE}
summary(affectedFig)
```
Above code is responsible for generating figure 3



#### Results


```{r echo=TRUE, tidy=TRUE}


#Plot Fatalities
 plot(affectedFig$ZONE, tothousehold/100000,  ylab="Total households in lakh", xlab = "Zone", main = "Total households zonewise for 2015 EQ") 

#add legends to graph
legend("topright", c("household", "Population", "Injured"), col=c("black", "red", "blue"), lty=1, lwd=1, bty="n")



library(plotrix)
slices <- totdeathbyzone$Tot_Deaths

colors <- rainbow(length(slices))
lbls <- totdeathbyzone$ZONE

par(mai = c(0,0,0,0))
layout(c(1,2),heights=c(0.6,1))
plot.new()
title(main="Pie Chart of death by zone ", line = -2)
legend("bottomright",legend=lbls, cex=0.8, fill=colors)

pie3D(slices,explode=0.2,
       col= colors )


```

### Summary

>From these data and graphs, we can further process for smart predictive analysis. Asking the right questions, using statistical inference, regression models for probability calcuation and other inference. 
